\ttitle

### 2013-11-05 Chapters/Chapter4/change_detection.tex
WIP: radius ratio thresholding

\begin{figure}
  \centering
    \includegraphics[width=\textwidth,height=\textheight,keepaspectratio]{./Figures/chapter4/expected_behaviour.pdf}
  \caption[Expected radius behaviour]{Expected behaviour of the radius $R$ of the hypersphere. The upper part shows a typical sinusoidal time series signal. The lower graph visualizes an abstract expectation of the values of $R$. Five possible windows are illustrated. The left, middle, and right windows cover an area of homogeneous signal. The expected value of $R$ is low. The other two windows cover a change entering and leaving the window, respectively. At these locations the radius $R$ is expected to increase temporarily. }
  \label{fig:radius_expectation}
\end{figure}

@@

\subsection{Ratio-thresholding}

\begin{equation}\label{eq:ratio_radius}
  \hbar_t = \frac{r_t}{\operatorname*{mean}(r_{y:t-1})},
\end{equation}
where $r_t$ is the approximated radius of the hypersphere at time $t$, $y$ is the time of the last change point (or $1$ if there is non) and $\operatorname*{mean}(r_{y:t-1})$ is the average of the previous approximated radii.

>>>>

\begin{figure}
  \centering
    \includegraphics[width=\textwidth,height=\textheight,keepaspectratio]{./Figures/chapter4/expected_behaviour.pdf}
  \caption[Expected radius behavior]{Expected behavior of the radius $R$ of the hypersphere. The upper part shows a typical sinusoidal time series signal. The lower graph visualizes an abstract expectation of the values of $R$. Five possible windows are illustrated. The left, middle, and right windows cover an area of homogeneous signal. The expected value of $R$ is low. The other two windows cover a change entering and leaving the window, respectively. At these locations the radius $R$ is expected to increase temporarily. }
  \label{fig:radius_expectation}
\end{figure}

@@

\subsection{Ratio-thresholding}
The first example of an algorithm that interprets the change indication time series and transforms it to change detection is the radius ratio thresholding method used by Camci~\cite{camci2010change}.
The method calculates the radius ratio $\hbar_t$ by taking the ratio of the current radius $r_t$ at time $t$ to the average radii since the last change point $y$:
\begin{equation}\label{eq:ratio_radius}
  \hbar_t = \frac{r_t}{\operatorname*{mean}(r_{y:t-1})},
\end{equation}
where $\operatorname*{mean}(r_{y:t-1})$ is the average of the previous approximated radii.
The ratio $\hbar_t$ is compared with two thresholds: the low threshold $th_\text{low}$ and the high threshold $th_\text{high}$.
When the data objects in the working set becomes more homogeneous at time $t$, the radius $R$ of the hypersphere and thereby the ratio $\hbar_t$ will decrease.
When the new value of $\hbar_r$ is lower than the low threshold $th_\text{low}$, a change detection is observed for time $t$.
The same holds for a more heterogeneous set of data and an increase of $\hbar_r$ higher than $th_\text{high}$.

The appropriate values for the thresholds are highly depended on the characteristics of the data and the trade-off parameter $C$ (as discussed in Section~\ref{subsec:oc-svm-svdd}).
This trade-off parameters regulates the fraction of data objects that will be rejected from the constructed model.
Since a high value of $C$ assigns a high penalty to outliers, data objects are more likely to be incorporated into the hypersphere by increasing the radius.
For low values of $C$ the cost of rejecting data objects is, compared to the benefits of a smaller hypersphere, relatively low.
This gives a relation between the parameters $C$, $th_\text{low}$, and $th_\text{high}$ and the sensitivity of the radius ratio based thresholding method.
A high value of $C$, or values close to $1$ for $th_\text{low}$ and $th_\text{high}$ result in a sensitive change detection procedure.
The reverse results in less sensitive methods.
The correct values for the intended sensitivity need to be empirically determined.
